---
title: "Lecture_notes"
author: "Jason_Z"
date: "5/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load data

```{r}
states = read.csv("http://bit.ly/dasi_states")


```



```{r}
#predicting poverty
pov_mlr = lm(poverty ~ female_house + white, data = states)
summary(pov_mlr)

```
```{r}
#testing the variability using ANOVA

anova(pov_mlr)



```

## Adjusted R^2

previously R^2 can be calculated by using formula R^2 = SSE/SST

Adjusted R^2 is taking account with only valuable variables.


```{r}
# R^2 adjusted = 1-(SSE/SST)*((n-1)/(n-k-1))

R_squared_adjusted <- 1-(339.49/480.25)*((51-1)/(51-2-1))

R_squared_adjusted
# compared to previously calculated as (480.25-339.49)/480.25 = 30.5%
# R^2 adjusted is always smaller than R^2
```
## definition of R-squared is to show the percentage of variability in the response variable explained by the model.
## definition of R-squared adjusted has a penalty for the number of predictor included in the model.



# Inference for multiple regression and model selection

load data 
```{r}
#load data

cognitive = read.csv("http://bit.ly/dasi_cognitive")

```

```{r}
#full model

cog_full = lm(kid_score~mom_hs + mom_iq + mom_work + mom_age, data = cognitive)

summary(cog_full)
```
'''
df = n-k-1
'''
df = 434 - 4 - 1 


F-statistics = 29.74

p-value: < 2.2e-16

p-value tells us the combination of the variables do not yield a good model